---
title: "Online Shopping"
author: "Griffin Kaiga"
date: "11/5/2020"
output: html_document
---

## Defining the question
The objective is to understand customers' behavior and learn the characteristics of various customer groups.

Defining the Metric for Success
the project will be considered a sucess once we are able to perform cluster analysis that will segment the customers into the various groups.

##Understanding the context
Kira Plastinina (Links to an external site.) is a Russian brand that is sold through a defunct chain of retail stores in Russia, Ukraine, Kazakhstan, Belarus, China, Philippines, and Armenia. The brand’s Sales and Marketing team would like to understand their customer’s behavior from data that they have collected over the past year. More specifically, they would like to learn the characteristics of customer groups.

## Experimental design
loading of data
cleaning the data
performing EDA
implementing the solution 
challenging the solution


```{r}
# reading the data
o.shop <- read.csv("shoppers.csv", header = TRUE)
head(o.shop)
View(o.shop)
```

```{r}
# checking the data types
str(o.shop)
```

```{r}
# checking for null values
is.na(o.shop)
# total number of missing values
colSums(is.na(o.shop))

# Dealing with missing values
o.shop <- na.omit(o.shop)

# check whether we still have missing values
anyNA(o.shop)
```
We had a total of 14 null row values for the columns Administrative, Administrative_Duration,




```{r}
library(dplyr)
# checking for outliers in our data
outliers <- boxplot(o.shop %>% select_if(is.numeric))
outliers

```


```{r}
# check for duplicates 
o.shop <- o.shop[!duplicated(o.shop),]
o.shop
```


```{r}
# checking for the unique values in each and every  column
lapply(o.shop, function(x)unique(x))
```


Univariate Analysis
```{r}
summary(o.shop)
```

```{r}
# mode of the columns
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

modes <- as.data.frame(c(getmode(o.shop$Month), getmode(o.shop$VisitorType), getmode(o.shop$Weekend), getmode(o.shop$Revenue)))
modes

```


```
```{r}
numeric.cols<- o.shop %>% select(1,2,3,4,5,6,7,8,9,10,12,13,14,15)



stddev = sapply(numeric.cols, sd)
quantile = sapply(numeric.cols, quantile)
stddev
quantile

```

```{r}
# frequency distribution for the categorical columns
categorical.cols <- o.shop %>% select(11, 16,17,18)

freq.Month <- as.data.frame(table(categorical.cols$Month))
freq.Month

freq.VisitorType <- as.data.frame(table(categorical.cols$VisitorType))
freq.VisitorType

freq.Weekend <- as.data.frame(table(categorical.cols$Weekend))
freq.Weekend

freq.Revenue <- as.data.frame(table(categorical.cols$Revenue))
freq.Revenue

```


```{r}
hist(o.shop$OperatingSystems)
hist(o.shop$Browser)
hist(o.shop$Region)
hist(o.shop$TrafficType)
```


Bivariate Analysis
```{r}
# checking for the correlation
library(ggcorrplot)
correlation = round(cor(select_if(o.shop, is.numeric)), 6)
ggcorrplot(corr, hc.order = T, ggtheme = ggplot2::theme_gray,
   colors = c("#6D9EC1", "white", "#E46726"), lab = T)
```

```{r}
# Scatter plot to for bounce rates and exit rates

plot(o.shop$BounceRates, o.shop$ExitRates, xlab="Bounce Rates", ylab="Exit Rates")
```


```{r}
# Comparing visitor type and revenue generation 

ggplot(data = o.shop) + 
  geom_bar(mapping = aes(x = VisitorType, fill = Revenue), position = "dodge")
```  
  



Implementing the solution

k-MEANS CLUSTERING

```{r}

```




```{r}
o.shop$Administrative_Duration[o.shop$Administrative_Duration<-1] <- 0

```
```{r}
# preprocessing
o.shop.new <- o.shop[, c(1,2,3,4,5,6,7,8,9,10,12,13,14,15)]
o.shop.class <- o.shop[, c(11,16,17,18)]
head(o.shop.new)

# previewing the class columns
head(o.shop.class)

```

```{r}
# Comparing revenue generation by region

ggplot(data = o.shop) + 
  geom_bar(mapping = aes(x = Month, fill = Revenue), position = "dodge")
```

```{r}

# Comparing revenue generation by special day status

ggplot(data = o.shop) + 
  geom_bar(mapping = aes(x = SpecialDay, fill = Revenue), position = "dodge")
```

```{r}
# converting categorical variables to numeric

o.shop$Month <- as.integer(as.factor(o.shop$Month))
o.shop$VisitorType <- as.integer(as.factor(o.shop$Month))
o.shop$Weekend <- as.integer(as.factor(o.shop$Weekend))
o.shop$Revenue <- as.integer(as.factor(o.shop$Revenue))
```

```{r}
# normalizing the dataset so that no particular atribute has more impact on the algorithm than others
normalize <- normalize <- function(x){
  return ((x-min(x)) / (max(x)-min(x)))
}

o.shop$Administrative <- normalize(o.shop$Administrative)
o.shop$Administrative_Duration <- normalize(o.shop$Administrative_Duration)
o.shop$Informational <- normalize(o.shop$Informational)
o.shop$Informational_Duration <- normalize(o.shop$Informational_Duration)
o.shop$ProductRelated <- normalize(o.shop$ProductRelated)
o.shop$ProductRelated_Duration <- normalize(o.shop$ProductRelated_Duration)
o.shop$BounceRates <- normalize(o.shop$BounceRates)
o.shop$ExitRates <- normalize(o.shop$ExitRates)
o.shop$PageValues <- normalize(o.shop$PageValues)
o.shop$SpecialDay <- normalize(o.shop$SpecialDay)
o.shop$OperatingSystems <- normalize(o.shop$OperatingSystems)
o.shop$Browser <- normalize(o.shop$Browser)
o.shop$Region <- normalize(o.shop$Region)
o.shop$TrafficType <- normalize(o.shop$TrafficType)
```
```{r}
# Performing clustering with a k value of 3

results = kmeans(o.shop, 3)

# Previewing the no. of records in each cluster

results$size
```

```{r}
# Getting the value of cluster center datapoint value(3 centers for k=3)
 
results$centers
```

```{r}
# Visualising the clusters 

plot(o.shop[2,3], col = results$cluster) 
```

```{r}
# Visualising the clusters 
plot(o.shop[c(3,4)], col = results$cluster)
plot(o.shop[c(5,6)], col = results$cluster)
plot(o.shop[c(1,2)], col = results$cluster)
plot(o.shop[c(7,8)], col = results$cluster)

table(results$cluster, o.shop$Revenue)

```

# Hierarchical Clustering
```{r}
# First we use the dist() function to compute the Euclidean distance between observations, 
# d will be the first argument in the hclust() function dissimilarity matrix
# ---
#
d <- dist(o.shop, method = "euclidean")

# We then hierarchical clustering using the Ward's method
# ---
# 
res.hc <- hclust(d, method = "ward.D2")
```

```{r}
#we plot the obtained dendrogram
# 
plot(res.hc, cex = 0.6, hang = -1)


```
